---
title: "Cars price data analysis"
author: "Mohammed Ansar Usman"
date: "6/30/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Open file
```{r open file}
df=read.csv("cars24data.csv",sep = ",")

```
## Structure of dataset
```{r Structure of Dataset}
str(df)

```
## Libraries
```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
library(forcats)
```


## Data preperation
```{r Libraries }
df=df %>% separate(t_model, into = c("year","brand","modelname"),sep = " ")

# Change the variable mileage in to numeric

df$t_mileage=gsub(",","",df$t_mileage)
df$t_mileage=gsub(" ","",df$t_mileage)
df$t_mileage=gsub("km","",df$t_mileage)
df$t_mileage=as.numeric(df$t_mileage)

# Variable t_cyl - change the variable in to two variable cyl and litre

df=df %>% separate(t_cyl, into = c("cyl","ltr"),sep=" ",fill = "right")

# Variable cylinder

df$cyl=gsub("cyl","",df$cyl)
df$cyl=ifelse(df$cyl=="1.5L",4,
              ifelse(df$cyl=="3.8L",6,df$cyl)
              )
df$cyl=as.numeric(df$cyl)

df$length=nchar(df$ltr)
df$ltr=ifelse(df$cyl==4 & nchar(df$ltr)==0,"1.5L",
              ifelse(df$cyl==6 & nchar(df$ltr)==0,"3.5L",
                     df$ltr))
df$ltr=gsub("L","",df$ltr,df$ltr)
df$ltr=as.numeric(df$ltr)

df=df %>% select(-X,-length)
df=df %>% distinct()

# Variable price

df$t_price=gsub("AED","",df$t_price)
df$t_price=gsub(",","",df$t_price)
df$t_price=as.numeric(df$t_price)

# Variable t_emi

df$t_emi=gsub("AED","",df$t_emi)
df$t_emi=gsub("/month","",df$t_emi)
df$t_emi=gsub(",","",df$t_emi)
df$t_emi=as.numeric(df$t_emi)

# Variable t_dp

df$t_dp=gsub("AED","",df$t_dp)
df$t_dp=gsub(",","",df$t_dp)
df$t_dp=gsub("downpayment","",df$t_dp)
df$t_dp=as.numeric(df$t_dp)

# Variable t_specs
df$option=ifelse(grepl("Basic Option",df$t_specs,ignore.case = T)==T,"Basic",
                 ifelse(grepl("Mid Option",df$t_specs,ignore.case = T)==T,"Mid",
                        ifelse(grepl("Full Option",df$t_specs,ignore.case = T)==T,"Full",df$t_specs)
                       )
                 )

anyNA(df)
lapply(df, anyNA)


df=df %>% mutate(Brand_Model=paste(brand,modelname))
t=sort(table(df$Brand_Model),decreasing = T)
head(t,30
     )
t=head(names(t),30)

df=df %>% select(-t_dp,-t_emi,-t_specs)


```
## Problem Statement
The main issue facing by used car sales companies is how the price will set based on market factors. 

The objective of this project is to scrape data from any prominent e-commerce website selling automobiles and analyze the factors correlated to price.

# Summary
```{r Summary of Dataset}

summary(df)
```


## Univariate analysis

### Year
```{r}

# Variable Year

df %>% ggplot(aes(year))+
  geom_bar(fill="royalblue",col="black")+
  labs(title ="Barplot of year",
       x="Year",
       y="Number of observations")


```
### Brand

```{r}

df %>% ggplot(aes(brand))+
  geom_bar(fill="cornflowerblue",col="black")+
  labs(title = "Barplot of Brand",
       x="Brand Name",
       y="Number of observations")+
  theme(axis.text.x = element_text(angle = 90,vjust = 0.5))

```

### Model Name

```{r}

df %>% group_by(modelname) %>%  
  summarise(No=n()) %>% 
  arrange(desc(No)) %>% 
  slice_max(No,n=30) %>% 
  ggplot(aes(modelname,No,fill=-No ))+
  geom_bar(stat = "identity")+
  theme(axis.text.x=element_text(angle = 90,vjust = 0.5))+
  labs(title ="Bar plot of top 30 Models",
       x="Model Name",
       y="No of observations")
  


```
### Mileage

```{r}
df %>% ggplot(aes(t_mileage))+
  geom_histogram(alpha=0.3,col="black",fill="royalblue",alpha=0.7)+
  labs(title = "Histogram of Mileage",
       x="Mileage",
       y="No.of observations")

xcut=cut(df$t_mileage,breaks = c(0,50000,100000,150000,200000))
options(scipen=999)
table(xcut)  
```
### Cylinder

```{r}
df %>% ggplot(aes(as.factor(cyl)))+
  geom_bar(fill="royalblue",col="black")+
  labs(title ="Barplot of cyl",
       x="Cylinder",
       y="Number of observations")

```

### Litre

```{r}
df %>% ggplot(aes(as.factor(ltr)))+
  geom_bar(fill="royalblue",col="black",alha=0.7)+
  labs(title ="Barplot of Litre",
       x="Litre",
       y="Number of observations")

```
### Price

```{r}
df %>% ggplot(aes(t_price))+
  geom_histogram(alpha=0.7,col="black",fill="royalblue")+
  labs(title = "Histogram of Price",
       x="Price",
       y='Number of observations')
options(scipen = 999)

xcut=cut(df$t_price,breaks = c(50000,75000,100000,125000,150000,200000))
table(xcut)

mean(df$t_price)
median(df$t_price)
sd(df$t_price)
sd(df$t_price)/mean(df$t_price)

```

## Bivariate Analysis
### Brand and Year

```{r}
df %>% ggplot(aes(brand,fill=year))+
  geom_bar()+
  scale_fill_brewer(palette = "Set1")+
  labs(title = "Brand and Year",
       x="Brand",
       y="No. of observations")+
  theme(axis.text.x = element_text(angle = 90))

# Stacked chart 

df %>% ggplot(aes(brand,fill=year))+
  geom_bar(position = "fill")+
  labs(title = "Brand and Year",
       x="Brand",
       y="No. of observations (%)")+
  
  scale_fill_brewer(palette = "Set1")+
  theme(axis.text.x = element_text(angle = 90))

  
```

### Model name and Year

```{r}



df %>%  filter(Brand_Model%in%t) %>% 
  ggplot(aes(fct_infreq(Brand_Model) ,fill=year))+
  geom_bar()+
  scale_fill_brewer(palette = "Set3")+
  theme(axis.text.x=element_text(angle = 90,vjust = 0.5))+
  labs(title = "Top 3O Models and Year",
       x="Model Name",
       y="No. of observations")
  
  
  
```
### Year and Mileage

```{r}
df %>% filter(t_mileage<=500000) %>% 
  ggplot(aes(year,t_mileage))+
  geom_boxplot()+
  geom_jitter(col="red",alpha=0.5,size=0.1)+
  stat_summary(fun = "mean",geom = "crossbar",col="blue")+
  labs(title = "Mileage and Model Year",
       subtitle = "Blue cross bar indicates mean of mileage",
       y="Mileage",
       x="Model Year")
```

### Option and Model Year

```{r}
df %>% ggplot(aes(year,fill=option))+
  geom_bar()+
  labs(title = " Model Year and Option",
       y="No. of observations",
       x="Model Year")


```
### Cylinder and Model Year
```{r}

df %>% ggplot(aes(year,fill=as.factor(cyl)))+
  geom_bar()+
  labs(title = "Cylinder and Model year",
       x="Cylinder",
       y="No. of observations")


```


### Brand and Mileage

```{r}
df %>%filter(t_mileage<=750000) %>%  
  ggplot(aes(brand,t_mileage))+
  geom_boxplot()+
  geom_jitter(size=0.2,col="red",alpha=0.5)+
  stat_summary(fun = "mean",geom="crossbar",col="blue")+
  theme(axis.text.x = element_text(angle = 90,vjust = 0.5))+
  labs(title = "Brand and Mileage",
       subtitle = "Blue crossbar indicates teh mean of mileage",
       x="Brand",
       y="Mileage")




```
### Coefficent of Mileage and Brand

```{r}

df %>%filter(t_mileage<=750000) %>%  
  group_by(brand) %>% 
  summarise(meanMileage=mean(t_mileage),
            medianMileage=median(t_mileage),
            sdMileage=sd(t_mileage),
            coe_sd_mileage=sdMileage/meanMileage,
            no=n()) %>% 
            arrange(desc(coe_sd_mileage)) %>% 
            ggplot(aes(no,coe_sd_mileage))+
  geom_point()+
  geom_text(aes(label=brand),size=2,vjust=3)+
  labs(title = "Coefficent of Mileage and Brand",
       x="No of car",
       y="Coefficent of variation - Mileage")

```

### Brand and Cylinder

```{r}
df %>% ggplot(aes(brand,fill=as.factor(cyl)))+
  geom_bar()+
  scale_fill_brewer(palette = "Set1")+
  theme(axis.text.x = element_text(angle = 90,vjust=0.5,hjust = 1))+
  labs(title = "Barplot of Brand and Cylinder",
       x="Brand",
       y="Number of observation")
```
### Brand and Price
```{r}
df %>%  
  ggplot(aes(brand,t_price))+
  geom_boxplot()+
  geom_jitter(size=0.2,col="red",alpha=0.5)+
  stat_summary(fun = "mean",geom="crossbar",col="blue")+
  theme(axis.text.x = element_text(angle = 90,vjust = 0.5))+
  labs(title = "Brand and Price",
       subtitle = "Blue crossbar indicates teh mean of price",
       x="Brand",
       y="Price")

```
### Brand and option
```{r}
df %>% ggplot(aes(brand,fill=option))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 90))
```


### Model name and Mileage

```{r}
df %>% filter(Brand_Model%in%t & t_mileage<=750000) %>% 
  ggplot(aes(Brand_Model,t_mileage))+
  geom_boxplot()+
  geom_jitter(size=0.2,col="red")+
  theme(axis.text.x = element_text(angle = 90,vjust = 0.5,hjust=1))+
  stat_summary(fun="mean",geom="crossbar",col="blue")+
  labs(title = "Model Name and Mileage",
       subtitle = "Jitter plot and box plot",
       x="Model Name",
       y="Mileage")

df %>% filter(Brand_Model%in%t & t_mileage<=750000) %>% 
  group_by(Brand_Model) %>% 
  summarise(meanMileage=mean(t_mileage),
          medianMileage=median(t_mileage),
          sdMileage=sd(t_mileage),
          coe_variation_mil=sdMileage/meanMileage,
          no=n()) %>% 
  ggplot(aes(no,coe_variation_mil))+
  geom_point()+
  geom_text(aes(label=Brand_Model),size=2,vjust=3)+
  labs(title = "Coefficent of Mileage and Model Name",
       x="No of car",
       y="Coefficent of variation - Mileage")




  
```
### Model Name and Price

```{r}
df %>% filter(Brand_Model%in%t) %>% 
  ggplot(aes(Brand_Model,t_price))+
  geom_boxplot()+
  geom_jitter(size=0.2,col="red")+
  theme(axis.text.x = element_text(angle = 90,vjust = 0.5,hjust=1))+
  labs(title = "Model Name and Price",
       subtitle = "Jitter plot and box plot",
       x="Model Name",
       y="Price")


df %>% filter(Brand_Model%in%t & t_mileage<=750000) %>% 
  group_by(Brand_Model) %>% 
  summarise(mean_price=mean(t_price),
          median_price=median(t_price),
          sd_price=sd(t_price),
          coe_variation_price=sd_price/mean_price,
          no=n()) %>% 
  ggplot(aes(no,coe_variation_price))+
  geom_point()+
  geom_text(aes(label=Brand_Model),size=2,vjust=3)+
  labs(title = "Coefficent of Price and Model Name",
       x="No of car",
       y="Coefficent of variation - Price")


```
### Modelname and option
```{r}
df %>% filter(Brand_Model%in%t) %>% 
  ggplot(aes(Brand_Model,fill=option))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 90,vjust = 0.5,hjust=1))
  
  

```


### Cylinder and Mileage

```{r}
df %>% filter(t_mileage<=750000) %>% 
  ggplot(aes(as.factor(cyl),t_mileage))+
  geom_boxplot()+
  geom_jitter(alpha=0.3,col="red",size=0.7)+
  stat_summary(geom="crossbar",col="blue",fun="mean")+
  labs(title = "Mileage and Cylinder",
       subtitle = "Blue line indicates the mean of Mileage",
       x="Cylinder",
       y="Mileage")


```
### Mileage and Price
```{r}
df %>% filter(t_mileage<=750000) %>% 
  ggplot(aes(t_mileage,t_price))+
  geom_point()+
  geom_smooth()+
  labs(title = "Mileage and Price",
       x="Price",
       y="Mileage")
  
  
```

### Cylinder and Price

```{r}
df %>% ggplot(aes(as.factor(cyl),t_price))+
  geom_boxplot()+
  geom_jitter(col="red",alpha=0.3,size=0.5)+
  labs(title = "Price and Cylinder",
       subtitle = "Blue bar indicates the mean of price",
       x="Cylinder",
       y="Price")
```
## Multivariate Analysis

### Price and Mileage based on Year

```{r}
df %>% filter(Brand_Model%in%t & t_mileage<=750000) %>% 
  ggplot(aes(t_price,t_mileage))+
  geom_point()+
  geom_smooth(method = lm)+
  facet_wrap(~year)

df %>% filter(Brand_Model%in%t & t_mileage<=750000) %>% 
  group_by(year) %>% 
  summarise(correlation=cor(t_price,t_mileage))

```

### Price and Mileage based on Brand

```{r}
df %>% filter(t_mileage<=750000) %>% 
  ggplot(aes(t_price,t_mileage))+
  geom_point()+
  geom_smooth(method = lm)+
  facet_wrap(~brand)

df %>% filter(t_mileage<=750000) %>% 
  group_by(brand) %>% 
  summarise(correlation=cor(t_price,t_mileage),
            no=n())

```

### Correlation of price and mileage based on Top 30 Model name, year and option
```{r}

df %>% filter(t_mileage<=750000 & Brand_Model%in%t) %>% 
  group_by(Brand_Model,year,option) %>% 
  summarise(Correlation=cor(t_mileage,t_price),No=n()) %>% 
  mutate(Model_year=paste(Brand_Model,year,option,sep = "_")) %>% 
  filter(Correlation<=-0.5) %>% 
  ggplot(aes(Model_year,Correlation,fill=No))+
  geom_bar(stat="identity")+
  theme(axis.text.x = element_text(angle = 90,hjust = 1,vjust = 0))+
  labs(title = "Correlation of Price and Mileage based on Top 30 Model,year and option",
       x="Model name, Year and Options",
       y="Correlation")+
  geom_text(aes(label=No),vjust=1)
  

  
```

```{r}


```


### Correlation of price and mileage based on Top 30 Model name


```{r}
df %>% filter(t_mileage<=750000 & Brand_Model%in%t) %>% 
  group_by(Brand_Model) %>% 
  summarise(Correlation=cor(t_mileage,t_price),No=n()) %>% 
  ggplot(aes(Brand_Model,Correlation,fill=No))+
  geom_bar(stat="identity")+
  theme(axis.text.x = element_text(angle = 90,hjust = 1,vjust = 0))+
  labs(title = "Correlation of Price and Mileage based on Model name",
       x="ModelName",
       y="Correlation")+
  geom_text(aes(label=No),vjust=1)

# Scatter plot

df %>% filter(Brand_Model%in%t & t_mileage<=750000) %>% 
  ggplot(aes(t_price,t_mileage))+
  geom_point()+
  geom_smooth(method = lm)+
  facet_wrap(~Brand_Model)


```


```{r}

df %>% filter(Brand_Model%in%t & t_mileage<=750000) %>% 
  ggplot(aes(t_price,t_mileage))+
  geom_point()+
  geom_smooth(method = lm)+
  facet_wrap(~cyl)


df %>% filter(Brand_Model%in%t & t_mileage<=750000) %>% 
  ggplot(aes(t_price,t_mileage))+
  geom_point()+
  geom_smooth(method = lm)+
  facet_wrap(~option)

df %>% filter(Brand_Model%in%t & t_mileage<=750000) %>% 
  ggplot(aes(t_price,t_mileage))+
  geom_point()+
  geom_smooth(method = lm)+
  facet_wrap(~brand)

df %>% filter(Brand_Model%in%t & t_mileage<=750000) %>% 
  ggplot(aes(t_price,t_mileage))+
  geom_point()+
  geom_smooth(method = lm)+
  facet_wrap(~ltr)

```


## Findings

+ Majority of observations are lying in year 2019.
+ Nissan, Mitsubishi and Toyota are the major brands based on observations.
+ Pajero, Patrol and Sunny are the major models based on observations.
+ In mileage, the majority of observations are lying in between 50000 to 100000.
+ 4 cylinder is listed more in e-commerce website.
+ 1.6L,2.0L and 3.5 is listed more in website.
+ Majority of price value is lying between 50000-75000 dirhams

+ 2019 model year car is more in top 3 brands ( Mitsubishi, Nissan and Toyota)
+ Nissan, Mitsubishi and Toyota are the major brands based on observations. The 2018 and 2019 model year cars are available more compare to other.
+ Pajero, Patrol and Sunny are the major models based on observations and 2019 model year is listed more.
+ The dispersion rate is high in 2016,2021 and 2022.
+ The mid and basic option is listed more.
+ The 4 cylinder and 6 cylinder cars are listed more in every model years
+ The 4 cylinder and 6 cylinder cars have noticed on top Brands like Nissan, Mitsubishi and Toyota
+ The mid and basic option is noticed on all brands and Models. The high option cars are very less.
+ Price and mileage have negative correlation ( -0.24 )
+ Price and Cylinder has positve correlation ( 0.58)




 







